---
title: "Exploratory analysis of Cnr1 in mouse Onecut3-GABAergic and dopaminergic hypothalamic neurons during development"
author: "Evgenii Tretiakov"
date: "`r Sys.Date()`"
output:
  workflowr::wflow_html:
    toc: true
---

```{r setup, include=FALSE, echo=FALSE}
docname <- "01-exploratory-data-analysis"
now <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    now <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - now)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.comments = FALSE,
  cache.lazy     = FALSE,
  dev            = c("png", "pdf"),
  echo           = TRUE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 14,
  fig.height     = 12,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
```
```{r load-packages}
# Load tidyverse infrastructure packages
library(here)
library(tidyverse)
library(magrittr)
library(zeallot)
library(future)

# Load packages for scRNA-seq analysis and visualisation
library(sctransform)
library(Seurat)
library(SeuratWrappers)
library(SeuratDisk)
library(scCustomize)
library(UpSetR)
library(patchwork)
library(RColorBrewer)
library(Nebulosa)
```
```{r set-paths}
src_dir <- here("code")
data_dir <- here("data")
output_dir <- here("output")
plots_dir <- here(output_dir, "figures")
tables_dir <- here(output_dir, "tables")
```
```{r load-functions}
source(here(src_dir, "functions.R"))
source(here(src_dir, "genes.R"))
```
```{r set-seed}
reseed <- 42
set.seed(seed = reseed)
```
```{r set-parallel-execution-plan}
# available cores
n_cores <- available_cores(prop2use = .5)
# Parameters for parallel execution
plan("multicore", workers = n_cores)
options(
  future.globals.maxSize = Inf,
  future.rng.onMisuse = "ignore"
)
plan()
```

## Read data
```{r load-data}
rar2020_ages_all <- c("E15", "E17", "P00", "P02", "P10", "P23")
rar2020_ages_postnat <- c("P02", "P10", "P23")
samples_df <- read_tsv(here("data/samples.tsv"))
colours_wtree <- setNames(
  read_lines(here(data_dir, "colours_wtree.tsv")),
  1:45
)

rar2020_srt_pub <-
  readr::read_rds(file.path(data_dir, "oldCCA_nae_srt.rds"))
rar2020_srt_pub %<>% UpdateSeuratObject()
colnames(rar2020_srt_pub@reductions$umap@cell.embeddings) <-
  c("UMAP_1", "UMAP_2")

rar2020_srt_pub$orig.ident <-
  rar2020_srt_pub %>%
  colnames() %>%
  str_split(pattern = ":", simplify = TRUE) %>%
  .[, 1] %>%
  plyr::mapvalues(
    x = .,
    from = samples_df$fullname,
    to = samples_df$sample
  )
rar2020_srt_pub$age <-
  plyr::mapvalues(
    x = rar2020_srt_pub$orig.ident,
    from = samples_df$sample,
    to = samples_df$age
  )
Idents(rar2020_srt_pub) <- "wtree"
neurons <-
  subset(rar2020_srt_pub, idents = c("18", "32"))
neurons <- RenameIdents(neurons, "18" = "PeVN_Onecut3", "32" = "Arc_TIDA")
neurons$celltype <- Idents(neurons)
neurons <-
  subset(neurons, subset = Slc17a6 == 0)
onecut3 <-
  subset(neurons,
    subset = Onecut3 > 0 | Th > 0 | Ddc > 0 | Slc6a3 > 0
  )

onecut3 <-
  Store_Palette_Seurat(
    seurat_object = onecut3,
    palette = rev(brewer.pal(n = 11, name = "Spectral")),
    palette_name = "div_Colour_Pal"
  )
```

Thus we subset the dataset to the neurons of interest from cluster 18, which are GABAergic and dopaminergic neurons expressing Onecut3 transcription factor. We also explicitly exclude Slc17a6-expressing neurons, which are glutamatergic neurons just in case to reduce noise. As the control group we use dopaminergic TIDA-neurons from the arcuate nucleus.


## Derive and filter matrix of neurons of interest
```{r derive-mtxs}
mtx_oc3 <-
  onecut3 %>%
  GetAssayData("data", "RNA") %>%
  as.data.frame() %>%
  t()
rownames(mtx_oc3) <- colnames(onecut3)

# Filter features
filt_low_genes <-
  colSums(mtx_oc3) %>%
  .[. > quantile(., 0.4)] %>%
  names()
mtx_oc3 %<>% .[, filt_low_genes]

min_filt_vector <-
  mtx_oc3 %>%
  as_tibble() %>%
  select(all_of(filt_low_genes)) %>%
  summarise(across(.fns = ~ quantile(.x, .1))) %>%
  as.list() %>%
  map(as.double) %>%
  simplify() %>%
  .[colnames(mtx_oc3)]

# Prepare table of intersection sets analysis
content_mtx_oc3 <-
  (mtx_oc3 > min_filt_vector) %>%
  as_tibble() %>%
  mutate_all(as.numeric)
```

## Plot UMAPs density

```{r plt-init-umap-density-onecut3, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Onecut3"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Th, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Th"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Slc6a3, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Slc6a3"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Prlr, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Prlr"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Cnr1, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Cnr1"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Gpr55, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Gpr55"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Mgll, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Mgll"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Dagla, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Dagla"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Daglb, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Daglb"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Faah, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Faah"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Napepld, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Napepld"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Gde1, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Gde1"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Pparg, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Pparg"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-slc32a1, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Slc32a1"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-gad1, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Gad1"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-gad2, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Custom(
  seurat_object = onecut3,
  features = c("Gad2"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

## Joint density UMAP's plots

```{r plt-init-umap-density-Onecut3-Cnr1, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Joint_Only(
  seurat_object = onecut3,
  features = c("Onecut3", "Cnr1"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Onecut3-Th, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Joint_Only(
  seurat_object = onecut3,
  features = c("Onecut3", "Th"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Cnr1-Th, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Joint_Only(
  seurat_object = onecut3,
  features = c("Cnr1", "Th"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Prlr-Cnr1, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Joint_Only(
  seurat_object = onecut3,
  features = c("Prlr", "Cnr1"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Prlr-Th, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Joint_Only(
  seurat_object = onecut3,
  features = c("Prlr", "Th"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Prlr-Onecut3, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Joint_Only(
  seurat_object = onecut3,
  features = c("Prlr", "Onecut3"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Gpr55-Prlr, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Joint_Only(
  seurat_object = onecut3,
  features = c("Gpr55", "Prlr"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Gpr55-Cnr1, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Joint_Only(
  seurat_object = onecut3,
  features = c("Gpr55", "Cnr1"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Gpr55-Th, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Joint_Only(
  seurat_object = onecut3,
  features = c("Gpr55", "Th"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```

```{r plt-init-umap-density-Gpr55-Onecut3, fig.width = 12, fig.height = 14, fig.fullwidth = TRUE}
Plot_Density_Joint_Only(
  seurat_object = onecut3,
  features = c("Gpr55", "Onecut3"),
  custom_palette = onecut3@misc$div_Colour_Pal
)
```


## Correlation analysis visualisation between different genes
```{r plot-correlation-scatter}
p_corrs <- list(
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    x = Onecut3,
    y = Cnr1,
    xfill = "#ffc400",
    yfill = "#e22ee2"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    x = Slc32a1,
    y = Onecut3,
    xfill = "#0000da",
    yfill = "#ffc400"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    x = Gpr55,
    y = Onecut3,
    xfill = "#006eff",
    yfill = "#ffc400"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    x = Slc32a1,
    y = Cnr1,
    xfill = "#0000da",
    yfill = "#e22ee2"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    x = Gpr55,
    y = Cnr1,
    xfill = "#006eff",
    yfill = "#e22ee2"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    x = Slc32a1,
    y = Gpr55,
    xfill = "#0000da",
    yfill = "#006eff"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    y = Slc32a1,
    x = Onecut3,
    yfill = "#0000da",
    xfill = "#ffc400"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    y = Gpr55,
    x = Onecut3,
    yfill = "#006eff",
    xfill = "#ffc400"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    y = Th,
    x = Onecut3,
    yfill = "#ff0000",
    xfill = "#ffc400"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    y = Gad1,
    x = Onecut3,
    yfill = "#a50202",
    xfill = "#ffc400"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    y = Gad2,
    x = Onecut3,
    yfill = "#4002a5",
    xfill = "#ffc400"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    y = Onecut2,
    x = Onecut3,
    yfill = "#6402a5",
    xfill = "#ffc400"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    y = Prlr,
    x = Onecut3,
    yfill = "#2502a5",
    xfill = "#ffc400"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    y = Ddc,
    x = Onecut3,
    yfill = "#4002a5",
    xfill = "#ffc400"
  ),
  ggstatsplot::ggscatterstats(
    as.data.frame(mtx_oc3),
    y = Slc6a3,
    x = Onecut3,
    yfill = "#2502a5",
    xfill = "#ffc400"
  )
)
n_corrs <- list(
  "oc3-rna-data-Onecut3-Cnr1",
  "oc3-rna-data-Slc32a1-Onecut3",
  "oc3-rna-data-Gpr55-Onecut3",
  "oc3-rna-data-Slc32a1-Cnr1",
  "oc3-rna-data-Gpr55-Cnr1",
  "oc3-rna-data-Slc32a1-Gpr55",
  "oc3-rna-data-Onecut3-Slc32a1",
  "oc3-rna-data-Onecut3-Gpr55",
  "oc3-rna-data-Onecut3-Th",
  "oc3-rna-data-Onecut3-Gad1",
  "oc3-rna-data-Onecut3-Gad2",
  "oc3-rna-data-Onecut3-Onecut2",
  "oc3-rna-data-Onecut3-Prlr",
  "oc3-rna-data-Onecut3-Ddc",
  "oc3-rna-data-Onecut3-Slc6a3"
)

walk2(n_corrs, p_corrs, save_my_plot, type = "stat-corr-plt")
```

### Visualise intersections sets that we are going to use (highlighted)

```{r plot-upset-oc3}
upset(
  as.data.frame(content_mtx_oc3),
  order.by = "freq",
  sets.x.label = "Number of cells",
  text.scale = c(2, 1.6, 2, 1.3, 2, 3),
  nsets = 15,
  sets = c("Th", "Onecut3"),
  queries = list(
    list(
      query = intersects,
      params = list("Onecut3"),
      active = T
    ),
    list(
      query = intersects,
      params = list("Th"),
      active = T
    ),
    list(
      query = intersects,
      params = list("Th", "Onecut3"),
      active = T
    )
  ),
  nintersects = 60,
  empty.intersections = "on"
)
```

```{r plot-upset-oc3-mod}
upset(
  as.data.frame(content_mtx_oc3),
  order.by = "freq",
  sets.x.label = "Number of cells",
  text.scale = c(2, 1.6, 2, 1.3, 2, 3),
  nsets = 15,
  sets = c("Gpr55", "Cnr1", "Slc32a1", "Th"),
  queries = list(
    list(
      query = intersects,
      params = list("Gpr55", "Slc32a1"),
      active = T
    ),
    list(
      query = intersects,
      params = list("Cnr1", "Th"),
      active = T
    )
  ),
  nintersects = 60,
  empty.intersections = "on"
)
```

Regroup factor by stages for more balanced groups
```{r derive-stages}
onecut3$age %>% forcats::fct_count()
onecut3$stage <-
  onecut3$age %>%
  forcats::fct_collapse(
    `Embrionic day 15` = "E15",
    `Embrionic day 17` = "E17",
    Neonatal = c("P00", "P02"),
    Postnatal = c("P10", "P23")
  )
onecut3$stage %>% forcats::fct_count()
```

Make subset of stable neurons
```{r derive-subset-def-cells}
onecut3$gaba_status <-
  content_mtx_oc3 %>%
  select(Gad1, Gad2, Slc32a1) %>%
  mutate(gaba = if_all(.fns = ~ .x > 0)) %>%
  .$gaba

onecut3$gaba_occurs <-
  content_mtx_oc3 %>%
  select(Gad1, Gad2, Slc32a1) %>%
  mutate(gaba = if_any(.fns = ~ .x > 0)) %>%
  .$gaba

onecut3$th_status <-
  content_mtx_oc3 %>%
  select(Th, Ddc, Slc6a3) %>%
  mutate(dopamin = if_any(.fns = ~ .x > 0)) %>%
  .$dopamin

oc3_fin <- onecut3
```

## Check contingency tables for neurotransmitter signature
```{r check-contingency-neurotransmitter-test}
oc3_fin@meta.data %>%
  janitor::tabyl(th_status, gaba_status)
```

By age
```{r check-contingency-neurotransmitter-age}
oc3_fin@meta.data %>%
  janitor::tabyl(age, th_status)
```

By stage
```{r check-contingency-neurotransmitter-stage}
oc3_fin@meta.data %>%
  janitor::tabyl(stage, th_status)
```

Make splits of neurons by neurotransmitter signature
```{r derive-subsets-by-neurotransmitter}
oc3_fin$status <- oc3_fin$th_status %>%
  if_else(true = "dopaminergic",
    false = "GABAergic"
  )
Idents(oc3_fin) <- "status"
SaveH5Seurat(
  object    = oc3_fin,
  filename  = here(data_dir, "oc3_fin"),
  overwrite = TRUE,
  verbose   = TRUE
)

## Split on basis of neurotrans and test for difference
oc3_fin_neurotrans <- SplitObject(oc3_fin, split.by = "status")

## Split on basis of age and test for difference
oc3_fin_ages <- SplitObject(oc3_fin, split.by = "age")
```

## DotPlots grouped by age

### Expression of GABA receptors in GABAergic Onecut3 positive cells
```{r plot-dotplot-age-gabar-gaba}
DotPlot(
  object = oc3_fin_neurotrans$GABAergic,
  features = gabar,
  group.by = "age",
  cols = c("#adffff", "#0084ff"),
  col.min = -1, col.max = 1
) + RotatedAxis()
```


### Expression of GABA receptors in dopaminergic TIDA or Onecut3 positive cells
```{r plot-dotplot-age-gabar-dop}
DotPlot(
  object = oc3_fin_neurotrans$dopaminergic,
  features = gabar,
  group.by = "age",
  split.by = "celltype",
  col.min = -1, col.max = 1
) + RotatedAxis()
```


### Expression of glutamate receptors in GABAergic Onecut3 positive cells
```{r plot-dotplot-age-glur-gaba, fig.height=5, fig.width=9}
DotPlot(
  object = oc3_fin_neurotrans$GABAergic,
  features = glutr,
  group.by = "age",
  cols = c("#adffff", "#0084ff"),
  col.min = -1, col.max = 1
) + RotatedAxis()
```

### Expression of glutamate receptors in dopaminergic TIDA or Onecut3 positive cells

```{r plot-dotplot-age-glur-dop, fig.height=7, fig.width=9}
DotPlot(
  object = oc3_fin_neurotrans$dopaminergic,
  features = glutr,
  group.by = "age",
  split.by = "celltype",
  col.min = -1, col.max = 1
) + RotatedAxis()
```

## DotPlots grouped by stage

### Expression of GABA receptors in GABAergic Onecut3 positive cells

```{r plot-dotplot-stage-gabar-gaba}
DotPlot(
  object = oc3_fin_neurotrans$GABAergic,
  features = gabar,
  group.by = "stage",
  cols = c("#adffff", "#0084ff"),
  col.min = -1, col.max = 1
) + RotatedAxis()
```


### Expression of GABA receptors in dopaminergic TIDA or Onecut3 positive cells

```{r plot-dotplot-stage-gabar-dop, fig.height=7, fig.width=9}
DotPlot(
  object = oc3_fin_neurotrans$dopaminergic,
  features = gabar,
  group.by = "stage",
  split.by = "celltype",
  col.min = -1, col.max = 1
) + RotatedAxis()
```


### Expression of glutamate receptors in GABAergic Onecut3 positive cells

```{r plot-dotplot-stage-glur-gaba, fig.height=5, fig.width=9}
DotPlot(
  object = oc3_fin_neurotrans$GABAergic,
  features = glutr,
  group.by = "stage",
  cols = c("#adffff", "#0084ff"),
  col.min = -1, col.max = 1
) + RotatedAxis()
```


### Expression of glutamate receptors in dopaminergic TIDA or Onecut3 positive cells

```{r plot-dotplot-stage-glur-dop, fig.height=7, fig.width=9}
DotPlot(
  object = oc3_fin_neurotrans$dopaminergic,
  features = glutr,
  group.by = "stage",
  split.by = "celltype",
  col.min = -1, col.max = 1
) + RotatedAxis()
```



### Expression of neuromodulators receptors in GABAergic Onecut3 positive cells

```{r plot-dotplot-stage-npr-gaba, fig.height=5, fig.width=15}
DotPlot(
  object = oc3_fin_neurotrans$GABAergic,
  features = npr,
  group.by = "stage",
  cols = c("#adffff", "#0084ff"),
  col.min = -1, col.max = 1
) + RotatedAxis()
```


### Expression of neuromodulators receptors in dopaminergic TIDA or Onecut3 positive cells

```{r plot-dotplot-stage-npr-dop, fig.height=6, fig.width=16}
DotPlot(
  object = oc3_fin_neurotrans$dopaminergic,
  features = npr,
  group.by = "stage",
  split.by = "celltype",
  col.min = -1, col.max = 1
) + RotatedAxis()
```


### Expression of endocannabinoids relevant genes in GABAergic Onecut3 positive cells

```{r plot-dotplot-stage-cnbn-gaba, fig.height=5, fig.width=9}
DotPlot(
  object = oc3_fin_neurotrans$GABAergic,
  features = cnbn,
  group.by = "stage",
  cols = c("#adffff", "#0084ff"),
  col.min = -1, col.max = 1
) + RotatedAxis()
```


### Expression of endocannabinoids relevant genes in dopaminergic TIDA or Onecut3 positive cells

```{r plot-dotplot-stage-cnbn-dop, fig.height=7, fig.width=9}
DotPlot(
  object = oc3_fin_neurotrans$dopaminergic,
  features = cnbn,
  group.by = "stage",
  split.by = "celltype",
  col.min = -1, col.max = 1
) + RotatedAxis()
```

## Overrepresentation analysis


```{r drv-sbs-mtx}
Idents(oc3_fin) <- "status"

sbs_mtx_oc <-
  oc3_fin %>%
  GetAssayData("data", "RNA") %>%
  as.data.frame() %>%
  t()
rownames(sbs_mtx_oc) <- colnames(oc3_fin)

# Filter features
filt_low_genes2 <-
  colSums(sbs_mtx_oc) %>%
  .[. > quantile(., 0.4)] %>%
  names()
sbs_mtx_oc %<>% .[, filt_low_genes2]

min_filt_vector2 <-
  sbs_mtx_oc %>%
  as_tibble() %>%
  select(all_of(filt_low_genes2)) %>%
  summarise(across(.fns = ~ quantile(.x, .005))) %>%
  as.list() %>%
  map(as.double) %>%
  simplify() %>%
  .[filt_low_genes2]

# Prepare table of intersection sets analysis
content_sbs_mtx_oc <-
  (sbs_mtx_oc > min_filt_vector2) %>%
  as_tibble() %>%
  mutate_all(as.numeric)
```

```{r plot-upset-sbs-oc3}
upset(
  as.data.frame(content_sbs_mtx_oc),
  order.by = "freq",
  sets.x.label = "Number of cells",
  text.scale = c(2, 1.6, 2, 1.3, 2, 3),
  nsets = 15,
  sets = c(
    "Th", "Onecut3",
    cnbn, "Prlr"
  ) %>%
    .[. %in% colnames(content_sbs_mtx_oc)],
  nintersects = 20,
  empty.intersections = NULL
)
```

```{r plot-upset-sbs-oc3-mod}
upset(
  as.data.frame(content_sbs_mtx_oc),
  order.by = "freq",
  sets.x.label = "Number of cells",
  text.scale = c(2, 1.6, 2, 1.3, 2, 3),
  nsets = 15,
  sets = c("Cnr1", "Prlr", "Th", "Onecut3") %>%
    .[. %in% colnames(content_sbs_mtx_oc)],
  nintersects = 10,
  empty.intersections = NULL
)
```

```{r plot-piecharts-sbs-oc3-chisqr, fig.height=12, fig.width=14}
sbs_mtx_oc_full <- content_sbs_mtx_oc |>
  select(any_of(c(
    neurotrans, cnbn, "Prlr", "Cnr1", "Gpr55", "Onecut3"
  ))) |>
  dplyr::bind_cols(oc3_fin@meta.data)

sbs_mtx_oc_full |> glimpse()

sbs_mtx_oc_full$CB1R <-
  sbs_mtx_oc_full %>%
  select(Cnr1, Gpr55) %>%
  mutate(CB1R = Cnr1 > 0) %>%
  .$CB1R

sbs_mtx_oc_full$PRLR <-
  sbs_mtx_oc_full %>%
  select(Prlr) %>%
  mutate(PRLR = Prlr > 0) %>%
  .$PRLR

sbs_mtx_oc_full$oc3 <-
  (sbs_mtx_oc_full$Onecut3 > 0)

library(ggstatsplot)
# for reproducibility
set.seed(123)

# plot
grouped_ggpiestats(
  # arguments relevant for `ggpiestats()`
  data = sbs_mtx_oc_full,
  x = CB1R,
  y = oc3,
  grouping.var = status,
  perc.k = 1,
  package = "ggsci",
  palette = "category10_d3",
  # arguments relevant for `combine_plots()`
  title.text = "Cnr1 specification of onecut-driven hypothalamic neuronal lineages by Onecut3 and main neurotransmitter expression",
  caption.text = "Asterisks denote results from proportion tests; \n***: p < 0.001, ns: non-significant",
  plotgrid.args = list(nrow = 2)
)

grouped_ggpiestats(
  # arguments relevant for `ggpiestats()`
  data = sbs_mtx_oc_full,
  x = PRLR,
  y = CB1R,
  grouping.var = status,
  perc.k = 1,
  package = "ggsci",
  palette = "category10_d3",
  # arguments relevant for `combine_plots()`
  title.text = "Prlr specification of onecut-driven PeVN or TIDA hypothalamic neuronal lineages by Cnr1 and main neurotransmitter expression",
  caption.text = "Asterisks denote results from proportion tests; \n***: p < 0.001, ns: non-significant",
  plotgrid.args = list(nrow = 2)
)

grouped_ggpiestats(
  # arguments relevant for `ggpiestats()`
  data = sbs_mtx_oc_full,
  x = PRLR,
  y = CB1R,
  grouping.var = celltype,
  perc.k = 1,
  package = "ggsci",
  palette = "category10_d3",
  # arguments relevant for `combine_plots()`
  title.text = "Prlr specification of onecut-driven PeVN or TIDA hypothalamic neuronal lineages by Cnr1",
  caption.text = "Asterisks denote results from proportion tests; \n***: p < 0.001, ns: non-significant",
  plotgrid.args = list(nrow = 2)
)
```
